### 2020-08-04 latex/vnhcs.tex
Input the first version of all VNHC theorems
This commit writes out the theorem stating the closed loop dynamics on
Gamma, as well as the corollary showing what happens when the system is
of degree of underactuation one.

Still to do is the proofs of these theorems, as well as the
characterization of the torque phase-feedback controller.

% TODO: Perform the full development of VNHCs, including its stabilizing
% controller.
\begin{defn}
    A \textit{virtual nonholonomic constraint} (VNHC) \textit{of order \(k\)} is a
    relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is
    smooth, \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
    \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
    stabilizing the set
    \[
        \Gamma = \left\{(q,p) \vert h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
    \]
    which is called the \textit{constraint manifold}.
\end{defn}

@@

% TODO: Write the [p' M p] matrix in our nice nottation
\begin{assm}\label{assm:H-is-simply-actuated}
    The mechanical system under consideration is an underactuated 
    Hamiltonian system of order \((n-k) \geq 0\) in simply
    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\):
    \begin{align*}
        \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q) \\
         &\begin{cases}
            \dot{q} = \Minv p \\
            \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q \Minv(q) p
            - \nabla_q V(q) + \simpleB \tau \\
        \end{cases}
    \end{align*}
\end{assm}

% TODO: write out ??? = [p' dM/(dqu_i)] the matrix of size nxn
\begin{prop}
    A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
    and
    \[
        \rank{\left(dh_q \Minv(q) + 
          dh_{p_u} (I_n \otimes p\tpose)\nabla_{q_u}\Minv(q) 
         \right)\simpleB} = k
    \]
    everywhere on the constraint manifold \(\Gamma\).
\end{prop}
\begin{proof}
\end{proof}

@@

>>>>

% TODO: Perform the full development of VNHCs, including its stabilizing
% controller.
\begin{defn}
    A \textit{virtual nonholonomic constraint} (VNHC) \textit{of order \(k\)} is a
    relation \(h(q,p) = 0\) where \(h : \mathcal{Q}\times\R^n \rightarrow \R^k\) is
    smooth, \(\rank{\left[ dh_q,\, dh_p \right]} = k\) for all 
    \((q,p) \in h\inv(0)\), and there exists a feedback controller \(\tau(q,p)\)
    stabilizing the set
    \[
        \Gamma = \left\{(q,p) \mid h(q,p) = 0, dh_q \dot{q} + dh_p \dot{p} = 0\right\}
    \]
    which is called the \textit{constraint manifold}.
\end{defn}

@@

% TODO: Write the [p' M p] matrix in our nice nottation
\begin{assm}\label{assm:H-is-simply-actuated}
    The mechanical system under consideration is a
    Hamiltonian system with \(n\) degrees of freedom and 
    \(k < n\) actuators. It is described in simply
    actuated coordinates \(\{q = (q_u,q_a), p = (p_u, p_a)\}\) and has the
    following dynamics:
    \begin{align*}
        \mathcal{H}(q,p) &= p\tpose \Minv(q) p + V(q) \\
         &\begin{cases}
            \dot{q} = \Minv p \\
            \dot{p} = -\frac{1}{2} (I_n \otimes p\tpose) \nabla_q \Minv(q) p
            - \nabla_q V(q) + \simpleB \tau \\
        \end{cases}
    \end{align*}
\end{assm}

\begin{thm}\label{thm:vnhc-regularity}
    A VNHC of order \(k\) is regular if and only if \(dh_{p_a} = 0\) 
    and
    \[
        \rank{\left(dh_q \Minv(q) + 
          dh_{p_u} (I_n \otimes p\tpose)\nabla_{q_u}\Minv(q) 
         \right)\simpleB} = k
    \]
    everywhere on the constraint manifold \(\Gamma\).
\end{thm}
\begin{proof}
    % TODO: complete this proof with our notation
\end{proof}

From the proof for Theorem \ref{thm:vnhc-regularity}, a regular VNHC of order
\(k\) can be stabilized by the output-linearizing phase-feedback controller
(\ref{eqn:vnhc-torque-controller}).
% TODO: Controller for VNHC in closed form
\begin{equation}\label{eqn:vnhc-torque-controller}
    \tau(q,p) = ? - k_p e - k_d \dot{e}
\end{equation}
where \(k_p, k_d \in \R_{>0}\) can be tuned using standard linear control
techniques.

Note that one generally cannot measure conjugate of momenta directly, as sensors
on mechanical systems typically only measure the state \((q,\dot{q})\). To
implement this controller in practice, one must compute \(p = M(q)\dot{q}\) at
every iteration. In other words, this controller requires knowledge of the full
state of the system.

@@

Now that we have found a controller to enforce a regular VNHC of order \(k\), we
would like to solve for the closed-loop dynamics. Naturally, these dynamics
should be specified through \((q_u, p_u)\) since \(q_a\) is a function of these
as specified by \(h(q,p_u) = 0\).
Unfortunately, \(\dot{q}_u\) depends on \(p_a\), and for general systems one
cannot solve explicitly for \(p_a\) in terms of \((q_u,p_u)\). This is because
the \(\dot{p}\) dynamics contains the coupling term 
\((I_n \otimes p\tpose)\nabla_{q_u}M(q)p\).

We now introduce a class of systems where explicitly solving for the closed-loop
dynamics is feasible.

\begin{defn}
    A mechanical system is ------ if \(\nabla_{q_u}M(q) = 0\)
\end{defn}

\begin{thm}\label{thm:zero-dynamics}
    Let \(\mathcal{H}\) be a ----- mechanical system satisfying Assumption
    \ref{assm:H-is-simply-actuated}. Let
    \(h(q,p_u) = 0\) be a regular VNHC of order \(k\) with constraint manifold
    \(\Gamma\).
    Then, on \(\Gamma\), the closed-loop dynamics are given by
    \begin{equation}\label{eqn:qpu-dynamics}
        \left.\begin{aligned}
            \dot{q}_u &= M(q)p \\
            \dot{p}_u &= -\nabla_{q_u}V(q) \\
            \end{aligned}\right|_{\begin{array}{c}
                h(q,p_u) = 0 \\ 
                pa = g(q_u,p_u) \\
            \end{array}}
    \end{equation}
    where
    % TODO: This is too wide, shrink it somehow
    \begin{equation}\label{eqn:g-qpu}
        g(q_u,p_u) := \left.\left(dh_q \Minv(q) \simpleB \right)\inv 
        \left(dh_{p_u} \nabla_{q_u}V(q) - dh_q \Minv(q)
        \begin{bmatrix}
            I_{n-k} \\
            \Zmat{k \times (n-k)} \\
        \end{bmatrix} p_u\right)\right|_{h(q,p_u) = 0}
    \end{equation}
\end{thm}
\begin{proof}
%TODO: Complete the proof for this theorem
\end{proof}

The most fascinating part of Theorem \ref{thm:zero-dynamics} is that it shows
that the dynamics on \(\Gamma\) are entirely described by the \(2(n-k)\)
unactuated coordinates, regardless of the number of degrees of freedom of the
system. This means that \(\Gamma\) is a \(2(n-k)\)-dimensional surface from
which the system cannot escape. This becomes especially important when analyzing
systems of degree of underactuation one, since it means the constraint is always
a curve which can be plotted on the plane with axes \(q_u\) and \(p_u\)
regardless of the original dimension \(n\) of the system.

In future sections of this thesis, we will apply a very specific type of
constraint. The following corollay applies Theorem \ref{thm:zero-dynamics} to
these constraints.

\begin{cor}\label{cor:2d-zero-dynamics}
    Suppose \(\mathcal{H}\) is a ---- mechanical system satisfying Assumption
    \ref{assm:H-is-simply-actuated} which has degree of underactuation one.
    Let \(h(q,p_u) = q_a - f(q_u,p_u)\) be a regular VNHC of order \((n-1)\),
    where \(f\) is a suitably-defined smooth function.
    Defining \(e_1 := (1,0,\ldots,0) \in \R^n\), the actuated momentum 
    is given by
    \begin{equation}\label{eqn:g-qupu}
        p_a = -\left.\left(dh_q \Minv(q)
        \begin{bmatrix}
            \Zmat{1\times (n-1)}\\
            I_{(n-1)}
        \end{bmatrix}\right)\inv 
        \left(\partial_{p_u}f \partial_{q_u}V + dh_q \Minv(q)e_1 p_u\right) 
            \right|_{q_a =f(q_u,p_u)}
    \end{equation}
    Since \(q_u \in \Rt{T}\) for some \(T > 0\), \((q_u(t),p_u(t))\) traces out
    a curve on the so-called \((q,p)\)-plane \(\Rt{T} \times \R\).
\end{cor}
